Millions of Type Errors

### 2023-09-01 programming23/paper.tex
region -> range, next step notes

{Roblox Studio} has a telemetry system, which is used to gauge
the effectiveness of creation features. This system stochastically
determines which sessions should report telemetry, and for those
sessions, reports telemetry records back with a summary of the
session. In the case of this study, the telemetry includes data on the
number of errors at various levels of granularity: in the current edit
region, in the current file, and in every file which was type
checked.

@@

\begin{description}
  \item[D1] metadata: uid, timestamp, session duration
  \item[D2] mode (\mnocheck{}, \mnonstrict{}, \mstrict{})
  \item[D3] a 1-bit reason for sending: did the user switch module?
    \subitem we get logs for two reasons: (1) switch module, (2) randomly with
    low probability after typechecking
  \item[D4] size of codebase: num files, num lines
  \item[D5] size of edit region
  \item[D6] overall type errors: for old and curr states, have: \# total, \# in module, and \# in region
  \item[D7] specific errors in the edit region: in the edit region, for old and
    curr states, how many type errors for each type error code? (There is 1
    edit region per log.)
    \subitem Examples, if the old state had one type-A error:
    \begin{itemize}
      \item If the latest edits touched it and fixed it, the next log says 1
        old type-A and 0 new type-A
      \item touched it, didn't fix => 1 old-A, 1 new-A
      \item didn't touch (fixed or not) => 0 old-A, 0 new-A
    \end{itemize}
  \item[D8] ForcedStrict "what if" errors for \mnocheck{}: specific errors for the
    edit region computed as if the user had strict mode enabled (even though they
    have no-check enabled)
    \subitem This info is available because it gives better hints for
    autocomplete, because it assumes precise types for ALL libraries (rather
    than Any)
  \item[D9] TooComplexErrors: project-wide count for this one specific error type
\end{description}

@@

\begin{itemize}
  \item How often does the typechecker give up with CodeTooComplex errors?
  \subitem [[ For context: the type system is designed to give useful feedback
    to all sorts of code, whether strict-typed, nonstrict-typed, or \mnocheck{}.
    But it has an escape hatch: CTC. Hope it's rarely needed. ]]
    \begin{itemize}
      \item Count the overall CTC errors (in each mode) for a coarse view of
        how rare they are. [D9 of course, D8 to compare \mnocheck{} code with
        strict-mode code via the same typechecker]
      \subitem Note: We can't look at the code to learn why the errors
        happened. Can only look for trends and tell the Roblox team where to
        look.
        %% \subitem Count CTC errors within each session to see if these errors are intermittent, or if they stick around. [D9 for the count, D8 for \mnocheck{} vs strict, D7 because CTC might show up in the edit region]
    \end{itemize}
  \item How helpful are highlighted regions?
    \begin{itemize}
      \item How often do edits in a highlighted region remove type errors? How
        often do they add new ones?
      \subitem [D7 compare specific old vs specific new]
      \subitem [may want to filter using D2 + D5]
      \item How often do edits outside the region remove type errors? How often do they add new ones?
      \subitem [D6 looking for old count vs. new count difference without a matching delta in the D7 counts]
      \item Which specific errors seem easy / hard to fix via highlights?
      \subitem [D6 removed by edit => easy; D6 survive edit => hard]
    \end{itemize}
  \item In what ways do types improve quality of life?
  \begin{itemize}
    \item Does \mnocheck{} code have many more latent errors than the others?
    \subitem [D8 latent errors come from the "what if" analysis]
  \end{itemize}
\end{itemize}

% What to study?
% 
% Sanity checks
% How large are edit regions? Plot the distribution.
% How long between logs? Long times & very short times are problematic.
% General Roblox trends
% How many sessions use strict / nonstrict / \mnocheck{}?
% Expect few strict/nonstrict, lots of \mnocheck{}
% How often does mode change within a session?
% Expect few of these b/c error reports are easy to ignore; non-intrusive
% If a down-switch happens, expect the current #errors to be high
% If an up-switch happens, hope to stay up
% How many CodeTooComplex errors arise, overall?
% Expect very few. Hopefully zero.
% How often do module switches happen?
% Expect not-too-many, with some logs in between each time
% Compare strict vs. nonstrict vs. nocheck
% Get median # errors across all snapshots.
% Expect nocheck median to be higher than the others.
% Plot #errors vs time, or vs codebase size
% Expect nocheck errors to keep going up. For other modes, #errors should stay low.
% General type error trends
% Plot the distribution of #errors across all edit regions
% Expect different counts based on error-kind. SEE BELOW
% Count specific errors that get fixed. For each kind of error, how many times is it old-in-region but not new-in-region? Compare to #old-in-region totals.
% Expect all errors to have a decent fix rate.
% If one error has a lower % than others, its error region may be misleading. Or, maybe it's hard to fix
% Type errors vs. edit region
% Compare #old vs. #new in region
% Expect a decrease, #old > #new
% If there's an increase, look for an explanation. Could be a huge edit region. Could be a 1-line region that we captured 2x in a row. Could be that this user is simply ignoring type errors.
% Compare [#old-in-region vs. (total #old - total #new)]
% If #old-in-region = 0, then hope the total # increases
% Decrease => nonlocal edits fixed an error => type error hilite may be inaccurate
% If #old-in-region > 0, then hope the total # decreases
% Increase (in region or otherwise) => made things worse

@@

\begin{figure}[t]
  %% TODO bigger text
  %% code/row-distribution.rkt
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded regions correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

\begin{figure}[t]
  TODO file size distribution: table or bars %% \includegraphics{img/row-distribution.pdf}
  \caption{Size of analyzed code: number of files, number of lines , and lines in edit region}
  \label{f:analysis-size}
\end{figure}

@@

FILL te in edit region, 
 14,917 non-stx
 16,007 syntax in edit regions.
 Most are typos!

@@

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll}
      \begin{tabular}[t]{lrl}
        595,137 & Type errors \\
        289,698 & in current module & [\pct{48.68}] \\
         30,924 & in edit region & [\pct{5.20}]
      \end{tabular}
      \begin{tabular}[t]{lrl}
        72,235,735 & {Forced-strict errors} \\
        37,027,281 & in current module & [\pct{51.26}] \\
         1,111,178 & in edit region & [\pct{1.54}]
      \end{tabular}
    \end{tabular}
    \caption{Type errors and forced-strict errors across all telemetry records}
    \label{f:total-tefs}
  \end{subfigure}

@@

@@

\begin{figure}[t]
  %% TODO stacked bars?
  %% TODO align old;curr
  \begin{tabular}{lrrrrr}
    Mode             & \# Rows & E         & Any Error & In Module & In Region \\
      & & & & (old ; curr) &  (old ; curr) \\\midrule
    \mnocheck{}   &   4,471 & \code{te} &    5.77\% &  0.20\% ;  2.21\% &  0.16\% ;  0.63\% \\
                  &         & \code{fs} &   59.18\% & 32.81\% ; 49.16\% &  9.17\% ; 13.13\% \\
    \mnonstrict{} &     506 & \code{te} &   25.89\% & 12.06\% ; 18.18\% &  6.32\% ;  7.91\% \\
                  &         & \code{fs} &   62.45\% & 37.94\% ; 53.75\% & 16.60\% ; 20.75\% \\
    \mstrict{}    &      24 & \code{te} &   50.00\% & 29.17\% ; 33.33\% &     0\% ;     0\% \\
                  &         & \code{fs} &   62.50\% & 50.00\% ; 58.33\% &  4.17\% ;  8.33\% 
  \end{tabular}

@@

  Below, edit region only.
  Each record with an edit region error adds +1 to at least 1 row.

@@

FILL error density: number vs region size

@@

Error counts per-project and per-module include syntax errors.
Problematic, because syntax errors are uninteresting (easy to fix),
but common, which makes it likely that our per-keystroke telemetry (1/2k) will record them.
Edit region has specific counts.
Next time, we should do that for the other aggregates.

>>>>

{Roblox Studio} has a telemetry system, which is used to gauge
the effectiveness of creation features. This system stochastically
determines which sessions should report telemetry, and for those
sessions, reports telemetry records back with a summary of the
session. In the case of this study, the telemetry includes data on the
number of errors at various levels of granularity: in the current edit
range, in the current file, and in every file which was type
checked.

@@

\begin{description}
  \item[D1] metadata: uid, timestamp, session duration
  \item[D2] mode (\mnocheck{}, \mnonstrict{}, \mstrict{})
  \item[D3] a 1-bit reason for sending: did the user switch module?
    \subitem we get logs for two reasons: (1) switch module, (2) randomly with
    low probability after typechecking
  \item[D4] size of codebase: num files, num lines
  \item[D5] size of edit range
  \item[D6] overall type errors: for old and curr states, have: \# total, \# in module, and \# in range
  \item[D7] specific errors in the edit range: in the edit range, for old and
    curr states, how many type errors for each type error code? (There is 1
    edit range per log.)
    \subitem Examples, if the old state had one type-A error:
    \begin{itemize}
      \item If the latest edits touched it and fixed it, the next log says 1
        old type-A and 0 new type-A
      \item touched it, didn't fix => 1 old-A, 1 new-A
      \item didn't touch (fixed or not) => 0 old-A, 0 new-A
    \end{itemize}
  \item[D8] ForcedStrict "what if" errors for \mnocheck{}: specific errors for the
    edit range computed as if the user had strict mode enabled (even though they
    have no-check enabled)
    \subitem This info is available because it gives better hints for
    autocomplete, because it assumes precise types for ALL libraries (rather
    than Any)
  \item[D9] TooComplexErrors: project-wide count for this one specific error type
\end{description}

@@

\begin{itemize}
  \item How often does the typechecker give up with CodeTooComplex errors?
  \subitem [[ For context: the type system is designed to give useful feedback
    to all sorts of code, whether strict-typed, nonstrict-typed, or \mnocheck{}.
    But it has an escape hatch: CTC. Hope it's rarely needed. ]]
    \begin{itemize}
      \item Count the overall CTC errors (in each mode) for a coarse view of
        how rare they are. [D9 of course, D8 to compare \mnocheck{} code with
        strict-mode code via the same typechecker]
      \subitem Note: We can't look at the code to learn why the errors
        happened. Can only look for trends and tell the Roblox team where to
        look.
        %% \subitem Count CTC errors within each session to see if these errors are intermittent, or if they stick around. [D9 for the count, D8 for \mnocheck{} vs strict, D7 because CTC might show up in the edit range]
    \end{itemize}
  \item How helpful are highlighted ranges?
    \begin{itemize}
      \item How often do edits in a highlighted range remove type errors? How
        often do they add new ones?
      \subitem [D7 compare specific old vs specific new]
      \subitem [may want to filter using D2 + D5]
      \item How often do edits outside the range remove type errors? How often do they add new ones?
      \subitem [D6 looking for old count vs. new count difference without a matching delta in the D7 counts]
      \item Which specific errors seem easy / hard to fix via highlights?
      \subitem [D6 removed by edit => easy; D6 survive edit => hard]
    \end{itemize}
  \item In what ways do types improve quality of life?
  \begin{itemize}
    \item Does \mnocheck{} code have many more latent errors than the others?
    \subitem [D8 latent errors come from the "what if" analysis]
  \end{itemize}
\end{itemize}

% What to study?
% 
% Sanity checks
% How large are edit ranges? Plot the distribution.
% How long between logs? Long times & very short times are problematic.
% General Roblox trends
% How many sessions use strict / nonstrict / \mnocheck{}?
% Expect few strict/nonstrict, lots of \mnocheck{}
% How often does mode change within a session?
% Expect few of these b/c error reports are easy to ignore; non-intrusive
% If a down-switch happens, expect the current #errors to be high
% If an up-switch happens, hope to stay up
% How many CodeTooComplex errors arise, overall?
% Expect very few. Hopefully zero.
% How often do module switches happen?
% Expect not-too-many, with some logs in between each time
% Compare strict vs. nonstrict vs. nocheck
% Get median # errors across all snapshots.
% Expect nocheck median to be higher than the others.
% Plot #errors vs time, or vs codebase size
% Expect nocheck errors to keep going up. For other modes, #errors should stay low.
% General type error trends
% Plot the distribution of #errors across all edit ranges
% Expect different counts based on error-kind. SEE BELOW
% Count specific errors that get fixed. For each kind of error, how many times is it old-in-range but not new-in-range? Compare to #old-in-range totals.
% Expect all errors to have a decent fix rate.
% If one error has a lower % than others, its error range may be misleading. Or, maybe it's hard to fix
% Type errors vs. edit range
% Compare #old vs. #new in range
% Expect a decrease, #old > #new
% If there's an increase, look for an explanation. Could be a huge edit range. Could be a 1-line range that we captured 2x in a row. Could be that this user is simply ignoring type errors.
% Compare [#old-in-range vs. (total #old - total #new)]
% If #old-in-range = 0, then hope the total # increases
% Decrease => nonlocal edits fixed an error => type error hilite may be inaccurate
% If #old-in-range > 0, then hope the total # decreases
% Increase (in range or otherwise) => made things worse

@@

\begin{figure}[t]
  %% TODO bigger text
  %% code/row-distribution.rkt
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded ranges correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

\begin{figure}[t]
  TODO file size distribution: table or bars %% \includegraphics{img/row-distribution.pdf}
  \caption{Size of analyzed code: number of files, number of lines , and lines in edit range}
  \label{f:analysis-size}
\end{figure}

@@

FILL te in edit range, 
 14,917 non-stx
 16,007 syntax in edit ranges.
 Most are typos!

@@

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll}
      \begin{tabular}[t]{lrl}
        595,137 & Type errors \\
        289,698 & in current module & [\pct{48.68}] \\
         30,924 & in edit range & [\pct{5.20}]
      \end{tabular}
      \begin{tabular}[t]{lrl}
        72,235,735 & {Forced-strict errors} \\
        37,027,281 & in current module & [\pct{51.26}] \\
         1,111,178 & in edit range & [\pct{1.54}]
      \end{tabular}
    \end{tabular}
    \caption{Type errors and forced-strict errors across all telemetry records}
    \label{f:total-tefs}
  \end{subfigure}

@@

%% TODO explain timestamps in prose, fig in appendix?

%% TODO power-law: event count, timespan, files lines editRange /session, 
%% ... gotta know project size

%% TODO overall stats on type errors

@@

\begin{figure}[t]
  %% TODO stacked bars?
  %% TODO align old;curr
  \begin{tabular}{lrrrrr}
    Mode             & \# Rows & E         & Any Error & In Module & In range \\
      & & & & (old ; curr) &  (old ; curr) \\\midrule
    \mnocheck{}   &   4,471 & \code{te} &    5.77\% &  0.20\% ;  2.21\% &  0.16\% ;  0.63\% \\
                  &         & \code{fs} &   59.18\% & 32.81\% ; 49.16\% &  9.17\% ; 13.13\% \\
    \mnonstrict{} &     506 & \code{te} &   25.89\% & 12.06\% ; 18.18\% &  6.32\% ;  7.91\% \\
                  &         & \code{fs} &   62.45\% & 37.94\% ; 53.75\% & 16.60\% ; 20.75\% \\
    \mstrict{}    &      24 & \code{te} &   50.00\% & 29.17\% ; 33.33\% &     0\% ;     0\% \\
                  &         & \code{fs} &   62.50\% & 50.00\% ; 58.33\% &  4.17\% ;  8.33\% 
  \end{tabular}

@@

  Below, edit range only.
  Each record with an edit range error adds +1 to at least 1 row.

@@

FILL error density: number vs range size

@@

Error counts per-project and per-module include syntax errors.
Problematic, because syntax errors are uninteresting (easy to fix),
but common, which makes it likely that our per-keystroke telemetry (1/2k) will record them.
Edit range has specific counts.
Next time, we should do that for the other aggregates.
